---
author: Cyrus Wakibia
title: Web Scraping Kenyan Weather
date: 04-09-2021
output:
  html_document:
    code_folding: hide
    toc: true
---

```{r setup, include=FALSE}
## load the libraries
library(dplyr)
library(rvest)
library(stringr)
library(knitr)
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
```

```{r, include=FALSE}
## site to scrape
url <- "https://meteo.go.ke/forecast/todays-weather"

## retrieve the url in R
weather_page <- read_html(url)

weather_page

```

## Web Scraping Kenya Meteorological Department

>- This tutorial shows how to scrape a weather website using r`rvest` package
>- The website can be accessed at [Kenya Meteorological Department](https://meteo.go.ke/forecast/todays-weather)
>- The towns whose wether have been included are only 24 but it's good to show how to scrape the data

## Web scraping procedure

>- You need to familiarise yourself with the **HTML and CSS language** which you can learn at [HTML Tutorials](https://www.w3schools.com/html/default.asp) and [CSS tutorials](https://www.w3schools.com/css/default.asp)
>- This will help you identifying the target nodes
>- Next you need to use the selectorGadget which is a google chrome extension and you can add it from [Selector gadget extension](https://chrome.google.com/webstore/detail/selectorgadget/mhjhnkcfbdhnjickkkdbjoemdmbfginb)
>- This is a video tutorial on how to use the selectorGadget [Video tutorial](https://selectorgadget.com/)
>- Finally you need to install the rvest package in R `install.packages("rvest")`
>- You can see some of the nodes and attributes among other things you can do to retrieve information from the site you have intended to scrape from [rvest tidyverse](https://rvest.tidyverse.org/articles/rvest.html)

## Variables we seek to get after performing web scraping

The variables we seek to get are:

>- Town
>- Weather Condition
>- Minimum Temperature
>- Maximum Temperature

We will finally create a data frame using these variables


## Perform Web Scraping on Kenya Meteorological Department


```{r}
# check the nodes in the site using the SelectorGadget
## town
town <- weather_page %>% 
  html_nodes(".views-field-field-town .field-content") %>% 
  html_text()

## weather condition
weather_condition <- weather_page %>% 
  html_nodes(".weather-desc") %>% 
  html_text()

## minimum temperature
min_temperature <- weather_page %>% 
  html_nodes(".temp-detail:nth-child(1)") %>% 
  html_text() %>% 
  ## extract only digits from the string
  str_extract("\\d.") %>% 
  ## convert the characters to strings
  as.numeric()

## maximum temperature
max_temperature <- weather_page %>% 
  html_nodes(".temp-detail+ .temp-detail") %>% 
  html_text() %>% 
  str_extract("\\d.") %>% 
  as.numeric()

```


## weather data frame

```{r}
## create the weather data frame
weather_data <- tibble(town, weather_condition,
                       ## temperature is in degree celcius
                       min_temperature, max_temperature)
kable(weather_data)
```

>- The above shows the data frame for the different weather conditions in 24 towns
>- Note that the site being scraped shows the most current weather conditions
